Viscous Transport in Eroding Porous Media

### 2019-10-01 docs/transport/erosion_transport.tex
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In equations~\eqref{eqn:velocityDLP} and~\eqref{eqn:deformationDLP},
$\uu$ and its deformation tensor are written as layer potentials in
$\RR^2$, but the quadrature method we introduce in
Section~\ref{sec:method} requires a complex-valued representation of the
layer potentials. The first step to form a complex representation is to
write the Laplace double-layer potential as the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Re (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \nick{ \CC(\Omega) },
  \label{eqn:laplaceComplex}
\end{align}
where $x,y \in \CC$ are the complex counterparts of $\xx,\yy \in \RR^2$,
and $\eta$ the complex counterpart of $\eeta$. \nick{We use the notation $\CC(\Omega)$ to denote the physical domain $\Omega$ represented in complex coordinates.}
\todo[inline]{What do you think of this notation I introduced to distinguish between $\Omega$ as a real versus complex domain? I was skimming over this section and was very confused about what $x$ and $y$ were, having naively assumed they were physical $x$ and $y$ coordinates instead of complex numbers. Hence I wanted to make their definitions more prominent. I realize that you state $x$ and $y$ are complex numbers, but it is in the text and could be missed by the careless reader (i.e. me).}
Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy integral by
first finding the boundary data of $v$.  Assuming that $\nick{\CC(\Omega)}$ is a
simply-connected interior domain, the boundary data of $v$ satisfies the
Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \nick{\CC(\bd\Omega)}.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Since $v$ is a holomorphic function and we
have its boundary data, by the Cauchy integral theorem we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \nick{\CC(\Omega)}$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

\todo[inline]{This figure seemed out of place to me because you do not present in other graphs yet, only the distribution of pores, and it is before the numerical methods are explained. HOWEVER, I think I see what you are driving towards: I think you want to give the reader a little teaser of the results, so that they are more inclined to make it through the methods section without losing interest (is that more or less right?). So perhaps just a few words to indicate our intent would prevent the reader from getting whiplash; maybe phrases like ``As illustration of the statistical results that will be elaborated in Section blah, we show here....''
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{sec:spatialDiscretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Since we use a BIE formulation, we only have to discretize the
one-dimensional boundary of the domain.  Therefore, we easily resolve
complex geometries, and differentiation is performed with spectral
accuracy in Fourier space.  We discretize each eroding body $\gamma_k$
with $N_\iin$ points and discretize the outer wall $\Gamma$ with
$N_\out$ points.  The $j^{th}$ discretization point on $\Gamma$ and
$\gamma_k$ are denoted by $\yy_j^0$ and $\yy^k_j$, respectively.  The
discretization points are initially distributed evenly in arclength, and
this equispacing is maintained for the entire simulation by using the
{\thL} formulation.  In addition, we apply mild smoothing
strategies~\cite{qua-moo2018} to avoid the corners that inevitably
develop in erosion.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Barycentric Quadrature Formulas}
\label{sec:bary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Since we are applying quadrature to periodic smooth functions, the
trapezoid rule is spectrally accurate~\cite{tre-wei2014}. However, since
the derivative of the integrands grows as the target point $\xx$
approaches $\bd\Omega$, the error bound of the trapezoid rule depends on
the location of the target point. Therefore, if the trapezoid rule, as
described in Section~\ref{sec:spatialDiscretization}, is applied when
bodies are in near-contact, or when a layer potential is evaluated at a
point close to $\bd\Omega$, then the results become unreliable and the
method ultimately becomes unstable.  Thus, we require a quadrature
method with an error bound that is independent of the target location.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this section, we apply our new quadrature method to simulate
erosion of several different configurations of nearly-touching eroding
bodies, and to characterize transport inside eroded geometries.  Each
eroding body is initially a circle with center $\cc_i$ and radius $r_i$.
The center and radius are chosen at random, and the body is accepted as
long as it is contained in $[-1,1] \times [-1,1]$ and is sufficiently
separated from all other bodies.  Since the error of our quadrature
method is bounded independent of the geometry, we consider bodies that
are separated by less than 10\% of an arclength spacing.  The randomized
method is repeated until the initial geometry reaches a desired initial
porosity.

@@

The erosion rate has a loss of regularity at stagnation points, and this
inevitably leads to corners developing on the eroding bodies.
Therefore, as described in Section~\ref{sec:formulation} and our
previous work~\cite{qua-moo2018}, we introduce a curvature penalization
term and a Gaussian smoothing step.  For all examples, we use the
smoothing parameters $\eps=15/256$ and $\sigma=10/256$, and the time
step size is $\Delta t = 10^{-4}$.

The common characteristic of each of the experiments is near-contact
between the eroding bodies, outer walls, and streamlines. We use our
numerical methods to describe and visualize the following examples:
\begin{itemize}
  \item{\bf Single Body Close to a Wall}: We consider a single eroding
  body close to the outer wall at $y=-1$.  We impose a shear flow
  centered at $y=-1$, we compare the eroding body's shape to a similar
  experiment of Spagnolie and Mitchell~\cite{mit-spa2017}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider a single eroding body close to a solid wall with the
shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
imposed on the outer wall. Mitchell and Spagnolie~\cite{mit-spa2017} did
a similar three-dimensional experiment using a second-order quadrature
method. Here, they initialized the eroding body so that its center $1.5$
times the body's radius above the solid wall.  In two dimensions, the
trapezoid rule is spectrally accurate, and an analogous initial
configuration is possible. However, unless the number of discretization
points is increased, the trapezoid rule becomes obsolete if the eroding
body is initialized closer to the solid wall. Therefore, to simulate
erosion near a solid wall at reasonable levels of discretization, we use
our hybrid method that combines the trapezoid rule and the Barycentric
quadrature rule.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In equations~\eqref{eqn:velocityDLP} and~\eqref{eqn:deformationDLP},
$\uu$ and its deformation tensor are written as layer potentials in
$\RR^2$, but the quadrature method we introduce in
Section~\ref{sec:method} requires a complex-valued representation of the
layer potentials. The first step to form a complex representation is to
write the Laplace double-layer potential as the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Re (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \nick{ \CC(\Omega) },
  \label{eqn:laplaceComplex}
\end{align}
where $x,y \in \CC$ are the complex counterparts of $\xx,\yy \in \RR^2$,
and $\eta$ the complex counterpart of $\eeta$. \nick{We use the notation $\CC(\Omega)$ to denote the physical domain $\Omega$ represented in complex coordinates.}
\todo[inline]{What do you think of this notation I introduced to distinguish between $\Omega$ as a real versus complex domain? I was skimming over this section and was very confused about what $x$ and $y$ were, having naively assumed they were physical $x$ and $y$ coordinates instead of complex numbers. Hence I wanted to make their definitions more prominent. I realize that you state $x$ and $y$ are complex numbers, but it is in the text and could be missed by the careless reader (i.e. me). However, I am not married to this notation and could easily be convinced just to leave it as you had it.}
Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy integral by
first finding the boundary data of $v$.  Assuming that $\nick{\CC(\Omega)}$ is a
simply-connected interior domain, the boundary data of $v$ satisfies the
Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \nick{\CC(\bd\Omega)}.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Since $v$ is a holomorphic function and we
have its boundary data, by the Cauchy integral theorem we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \nick{\CC(\Omega)}$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

\todo[inline]{This figure seemed out of place to me because you do not present other graphs yet, only the distribution of pores, and it is before the numerical methods are explained. HOWEVER, I think I see what you are driving towards: I think you want to give the reader a little teaser of the results, so that they are more inclined to make it through the methods section without losing interest (is that more or less right?). So perhaps just a few words to indicate our intent would prevent the reader from getting whiplash; maybe phrases like ``As illustration of the statistical results that will be elaborated in Section blah, we show here....''
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{sec:spatialDiscretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Since we use a BIE formulation, we only have to discretize the
one-dimensional boundary of the domain.  Therefore, we easily resolve
complex geometries, and differentiation is performed with spectral
accuracy in Fourier space.  We discretize each eroding body $\gamma_k$
with $N_\iin$ points and discretize the outer wall $\Gamma$ with
$N_\out$ points.  The $j^{th}$ discretization point on $\Gamma$ and
$\gamma_k$ are denoted by $\yy_j^0$ and $\yy^k_j$, respectively.  The
discretization points are initially distributed evenly in arclength, and
this equispacing is maintained for the entire simulation by using the
{\thL} formulation.  In addition, we apply regularization~\cite{qua-moo2018} to \nick{\st{avoid} slightly smooth} the corners that inevitably develop during erosion.
\todo[inline]{Note: We do not want to say that we avoid corners because that makes it sound like our numerical simulation fails to capture the real physical phenomenon of corner formation.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Barycentric Quadrature Formulas}
\label{sec:bary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\st{Since we are applying quadrature to periodic smooth functions, the
trapezoid rule is spectrally accurate. However, since
the derivative of the integrands grows as the target point $\xx$
approaches $\bd\Omega$, the error bound of the trapezoid rule depends on
the location of the target point.}
\nick{While the trapezoid rule is spectrally accurate for smooth, periodic functions~\cite{tre-wei2014}, the derivative of the integrands grows as the target point $\xx$ approaches $\bd\Omega$ and so the error bound depends on the location of the target point.} Therefore, if the trapezoid rule is applied when bodies are in near-contact, or when a layer potential is evaluated at a point close to $\bd\Omega$, then the results become unreliable and the \nick{simulation} ultimately becomes unstable.  We thus \nick{desire a quadrature method whose error bound does not depend on the target location}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\nick{In this section, we apply our new quadrature method to simulate
erosion of several different configurations of nearly-touching eroding
bodies, and to characterize transport inside eroded geometries.}
\todo[inline]{This opening sentence is awkward and seems like a run-on sentence.}
Each eroding body is initially a circle with center $\cc_i$ and radius $r_i$.
The center and radius are chosen at random, and the body is accepted \nick{if} it is contained in $[-1,1] \times [-1,1]$ and is sufficiently
separated from all other bodies.  Since the error of our quadrature
method is bounded independent of the geometry, we consider bodies that
are separated by less than 10\% of an arclength spacing.  The randomized
method is repeated until the initial geometry reaches a desired initial
porosity.

@@

The erosion rate \nick{\st{has a loss of regularity} loses regularity at stagnation points, which inevitably leads to corner formation on the bodies.} 
As described in Section~\ref{sec:formulation} and our
previous work~\cite{qua-moo2018}, we introduce a curvature penalization
term and a Gaussian smoothing step in order to ameliorate corner formation.  
For all examples, we use the smoothing parameters $\eps=15/256$ and $\sigma=10/256$, and the time
step size is $\Delta t = 10^{-4}$.

The common characteristic of each of the experiments is near-contact
between the eroding bodies, outer walls, and streamlines. We use our
numerical methods to describe and visualize the following examples:
\begin{itemize}
  \item{\bf Single Body Close to a Wall}: We consider a single eroding
  body close to the outer wall at $y=-1$.  We impose a shear flow
  centered at $y=-1$ \nick{and} compare the eroding body's shape to a similar
  experiment of Spagnolie and Mitchell~\cite{mit-spa2017}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\nick{\st{We} First,} consider a single eroding body close to a solid wall with the
shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
imposed on the outer wall. Mitchell and Spagnolie~\cite{mit-spa2017} \nick{\st{did} performed} a similar three-dimensional experiment using a second-order quadrature
method. \st{Here} They initialized the \st{eroding} body \st{so that} with its center located $1.5$ \nick{radii} above the solid wall.  In two dimensions, the
trapezoid rule is spectrally accurate, and an analogous initial
configuration is possible. However, unless the number of discretization
points is increased, the trapezoid rule becomes obsolete if the eroding
body is initialized closer to the solid wall. Therefore, to simulate
erosion near a solid wall at reasonable levels of discretization, we use
our hybrid method that combines the trapezoid rule and the Barycentric
quadrature rule.
